---
title: "Result analysis"
output:
  html_document:
    df_print: paged
    toc_depth: '2'
  html_notebook: default
  pdf_document:
    highlight: zenburn
    keep_tex: yes
params:
  run_test: no
  use_kable: no
---

```{r setup, include=FALSE}

library(knitr)
library(ggsci)
library(tidyverse)
library(glue)
library(cowplot)
library(ggrepel)

source("functions_learning.R")
source("init.R")

knitr::opts_chunk$set(
  echo = FALSE, 
  message = FALSE, 
  warning = FALSE,
  fig.width=fig.width.baseline*output.size, 
  out.width=out.width.default)

theme_set(theme_bw(base_size = 9.5) + theme(legend.title=element_blank()))

```


We use this file to compare the final results from analysis_learningrates_corrections.

```{r load all results}

all_learnig_rates_results <- list()
all_learnig_rates_average_costs <- list()

for(currency in relevant_currencies){
  all_learnig_rates_results[[paste(currency,"real")]]  <- read_csv(paste0("output/results/learning_rate_results_",currency,"_real.csv"))
  all_learnig_rates_results[[paste(currency,"nominal")]]  <- read_csv(paste0("output/results/learning_rate_results_",currency,"_nominal.csv"))
  all_learnig_rates_average_costs[[paste(currency,"real")]]  <- read_csv(paste0("output/results/learning_rate_average_costs_",currency,"_real.csv"))
  all_learnig_rates_average_costs[[paste(currency,"nominal")]]  <- read_csv(paste0("output/results/learning_rate_average_costs_",currency,"_nominal.csv"))
}


all_results <- all_learnig_rates_results %>% 
  reduce(bind_rows) %>% 
  distinct() %>% 
  mutate(
    type = fct_relevel(type, "uncorrected_project_weighted", "uncorrected_marketshare_weighted", "corrected_marketshare_weighted"),
    type_label = type,
    type_label = fct_recode(type_label,
 "Uncorrected \n project-wt." = "uncorrected_project_weighted",
 "Uncorrected \n market share-wt." = "uncorrected_marketshare_weighted",
 "Fx-corrected \n market share-wt." = "corrected_marketshare_weighted"),
    interval = fct_relevel(interval, factor_order_intervals),
    interval = fct_recode(interval,!!!plot_stips_label)
    ) %>% 
  arrange(currency, interval)


all_costs <- all_learnig_rates_average_costs %>% 
  reduce(bind_rows) %>% 
  distinct() %>% 
  mutate(
    type = fct_relevel(type, "uncorrected_project_weighted", "uncorrected_marketshare_weighted", "corrected_marketshare_weighted"),
    type_label = type,
    type_label = fct_recode(type_label,
 "Uncorrected \n project-wt." = "uncorrected_project_weighted",
 "Uncorrected \n market share-wt." = "uncorrected_marketshare_weighted",
 "Fx-corrected \n market share-wt." = "corrected_marketshare_weighted")
    ) %>% 
  arrange(currency, year)


cum_capacity <- read_csv(paste0("output/results/learning_rate_cum_cap.csv"))


```


# Plots

## Combined plot of uncorrected

To plot all learning rates in the same plot, it is necessary to norm the costs. To do this, we select the initial year's average global costs and divide all average global costs by that value. We then get 1/W for all average costs and can compare the learning rates.

```{r Combined and normed plot, echo=FALSE, fig.cap="Comparison of learning rates, as index of the average costs in 2006", fig.width=fig.width.baseline*output.size, fig.asp=0.4, out.width=out.width.default}


combined_plot_data <- all_costs %>% 
  filter(
    type == "uncorrected_marketshare_weighted",
    version == "real"
  ) %>% 
  inner_join(cum_capacity) %>% 
  filter(year >= T0, year <= T_max) %>% 
  select(
    year, 
    average_costs, 
    currency, 
    cumulative_capacity = X) %>% 
  mutate(currency = factor(currency, plot_order_currency))

combined_plot_norm_data <- combined_plot_data %>% 
  filter(year == T0) %>% 
  select(norm_costs = average_costs, currency)
  
normed_plot_data <- combined_plot_data %>% 
  inner_join(combined_plot_norm_data, by = "currency") %>% 
  mutate(normed_average_costs = average_costs / norm_costs) %>% 
  arrange(currency, year)

normed_plot_data_subsets <- list()
for(i in seq_along(intervals)){
  interval <- intervals[[i]]
  interval_name <- names(intervals[i])
  interval_years <- as.numeric(interval[1]):as.numeric(interval[2])
  
  x_for_abclabel <- normed_plot_data %>% 
    filter(year %in% interval_years[[1]]) %>% 
    pull(cumulative_capacity)
  x_for_abclabel <- x_for_abclabel[[1]]/1000
  
  normed_plot_data_subsets[[i]] <- normed_plot_data %>% 
    filter(year %in% interval_years) %>% 
    add_column(interval = interval_name) %>% 
    add_column(x_for_abclabel = !!x_for_abclabel)
}

normed_plot_data_reduced <- normed_plot_data_subsets %>% 
  reduce(bind_rows) %>% 
  mutate(
    interval = fct_relevel(interval, names(intervals)),
    interval = fct_recode(interval,!!!plot_stips_label))

plot_labels <- normed_plot_data_reduced %>% 
  expand(interval) %>% 
  add_column(label = letters[seq( from = 1, to = 3 )])

normed_plot_data_reduced %>% 
  mutate(year_for_text = if_else(currency == "GBP", str_sub(as.character(year), start = 3, end = 4) , "")) %>%
  ggplot(aes(cumulative_capacity/1000, normed_average_costs)) +
    geom_smooth(
      aes(col = currency),
      method="lm", formula = (y ~ x), se=FALSE) +
    geom_point(aes(col = currency)) +
    geom_text_repel(aes(label = year_for_text), vjust = -0.6, hjust = 0.1, size = 3, point.padding = NA) +
    geom_text(data = plot_labels, mapping = aes(x = c(250,60,250), y = c(1.2,1.2,1.2), label = label), fontface = "bold") +

    scale_x_continuous(trans="log", breaks = c(c(1,seq(2,10,2)) %o% 10^(0:4)), minor_breaks = 0.5) +
    scale_y_continuous(trans="log", breaks = c(seq(0,1,0.1), seq(1.2,2,0.2)), minor_breaks = NULL) +
    theme(legend.position="bottom") +
    guides(
      linetype = guide_legend(title="Learning rate", nrow = 1), 
      col = guide_legend(title="Learning rate", nrow = 1), 
      size = guide_legend(title="Type")) +
    labs(x = "Cumulative capacity [GW]", 
         y = paste0("Index of costs per MW")) +
    scale_color_manual(values = plot_colours_currencies) +
  facet_grid( ~ interval, scales = "free_x", space = "free_x")


```
\clearpage


As asked by reviewer:

```{r, fig.asp=0.4, out.width=out.width.default}

normed_plot_data_reduced %>% 
  mutate(year_for_text = if_else(currency == "GBP", str_sub(as.character(year), start = 3, end = 4) , "")) %>%
  ggplot(aes(cumulative_capacity/1000, normed_average_costs)) +
    geom_smooth(
      aes(col = currency, linetype = interval),
      method="lm", formula = (y ~ x), se=FALSE) +
    #geom_point(aes(col = currency)) +
    #geom_text_repel(aes(label = year_for_text), vjust = -0.6, hjust = 0.1, size = 3, point.padding = NA) +
    #geom_text(data = plot_labels, mapping = aes(x = c(250,60,250), y = c(1.2,1.2,1.2), label = label), fontface = "bold") +

    scale_x_continuous(trans="log", breaks = c(c(1,seq(2,10,2)) %o% 10^(0:4)), minor_breaks = 0.5) +
    scale_y_continuous(trans="log", breaks = c(seq(0,1,0.1), seq(1.2,2,0.2)), minor_breaks = NULL) +
    theme(legend.position="right") +
    # guides(
    #   linetype = guide_legend(title="Learning rate", nrow = 1), 
    #   col = guide_legend(title="Learning rate", nrow = 1), 
    #   size = guide_legend(title="Type")) +
    labs(x = "Cumulative capacity [GW]", 
         y = paste0("Index of costs per MW"),
         linetype = "Interval",
         col = "Currency") +
    scale_color_manual(values = plot_colours_currencies) #+
  #facet_grid( ~ , scales = "free_x", space = "free_x")


```


```{r all_results_plot, fig.cap="Learning rates plot", fig.asp=0.4}


plot_labels = all_results %>% 
  expand(version, interval) %>% 
  add_column(label = letters[seq( from = 1, to = 6 )])
  
all_learnig_rates_results_plot = all_results %>% 
  ggplot(mapping = aes(x = type_label, y = estimate, fill = currency)) +
    geom_col(position = "dodge2") +
    geom_text(. %>% filter(type_label == type_label), mapping = aes(y = estimate/2, label = round(estimate)), 
              position = position_dodge(0.9), col = "white", size = 2.8, angle = 90) +
    geom_text(data = plot_labels, mapping = aes(x = 0.6, y = 43, label = label), fontface = "bold") +
    scale_y_continuous(limits = c(0,44), expand = expand_scale(mult = c(0, 0.1))) +
    labs(
      y = paste0("Learning rate in %")) +
    theme(
      axis.title.x = element_blank(),
      panel.grid.major = element_blank(),
      legend.position="bottom",
      legend.spacing.x = unit(0.1, 'cm')) +
    guides(
      fill = guide_legend(nrow = 1)) +
    facet_grid(version ~ interval, switch = "y") +
    scale_fill_manual(values = plot_colours_currencies)

all_learnig_rates_results_plot


```


## Combined plot

Comparison with the simple learning rates for every country

```{r Combined plot, fig.cap="Learning rates plot", fig.asp=0.4}

#mypal = pal_npg("nrc")(length(relevant_currencies))

all_results_for_combined_plot <- all_results %>% 
  filter(version == "real")

# comparision currency (USD)
corrected_rates <- all_results_for_combined_plot %>% 
  filter(
    currency == "USD",
    type == "corrected_marketshare_weighted") %>% 
  select(corrected_lr = estimate, interval)


bind_rows(
  all_results_for_combined_plot %>% 
    filter(
      currency == "USD",
      type == "corrected_marketshare_weighted") %>% 
    add_column(legend = factor("fx-corrected (USD)")),
  all_results_for_combined_plot %>% 
    filter(
      type == "uncorrected_marketshare_weighted") %>% 
    mutate(legend = currency)
  ) %>%  
  get_comparison_plot(corrected_rates)


```







# Summary



```{r summary table}

all_types <- all_results$type %>% levels()
all_intervals <- all_results$interval %>% levels()

library(kableExtra)

tables <- list()

# Subtitle per step 1,2,3
for(type in all_types){
  
  tables[[type]] <- bind_cols(
     filter(all_results, type == !!type, interval == all_intervals[1]),
     filter(all_results, type == !!type, interval == all_intervals[2]),
     filter(all_results, type == !!type, interval == all_intervals[3])
  ) %>% 
    arrange(version, currency) %>% 
    select(currency, estimate, CI, R2, estimate1, CI1, R21,estimate2, CI2, R22)
    
  kable_table <- tables[[type]] %>% 
    kable(caption = glue("Results for {type}.")) %>%
    kable_styling() %>% 
    add_header_above(c(" " =1, 
                       "Full 10-year interval (2006-2016)" = 3, 
                       "First 5-year interval (2006-2011)" = 3, 
                       "Second 5-year interval (2011-2016)" = 3)) %>% 
    kableExtra::group_rows("nominal", 1, 6) %>%
    kableExtra::group_rows("real", 7, 12)
  
  print(kable_table)
  
  write_excel_csv(
    tables[[type]],
    glue("output/reports_tables/results_table_{type}.csv"))
  
}


```